Unknown

### 2024-04-03 ms/ms.qmd
finished drafting data cleaning methods

library(readr)
# I am mostly putting placeholders here for objects/values that will need to be
# loaded eventually to use in paper.
aa_args = read_rds("../objects/aa_args.rds")
fit_aa = read_rds("../objects/fit_dummy.rds")
dim_fit_aa = dim(fit_aa$fit)

@@

### Cleaning RH curves
n_rh_curve_raw = 2400 # export from adaptive-amphistomy/r/07_process-licor.R
n_points_per_rh_curves_raw = 100 # export from adaptive-amphistomy/r/07_process-licor.R
ca_min = aa_args$ca_min
ca_max = aa_args$ca_max
n_min_points_per_rh_curve = aa_args$n_min_points_per_rh_curve
outlier_threshold = aa_args$outlier_threshold

@@

1. remove unreliable data points
2. remove hysteretic portion of \agcurve{} curves at low \gsw{}
3. remove outliers within each \agcurve{} curve
4. remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves
5. remove replicates with outlier \aax values

@@

*Procedure*: We fit provisional quadratic regressions for each curve using ordinary least squares with the `lm()` function in *R*. We sequentially removed data points with an external studentized residual greater than `r outlier_threshold` until none remained.

were likely due to measurement error or leaf damage.
WORKING HERE

>>>>

library(readr)
# I am mostly putting placeholders here for objects/values that will need to be
# loaded eventually to use in paper.
aa_args = read_rds("../objects/aa_args1.rds")
fit_aa = read_rds("../objects/fit_dummy.rds")
dim_fit_aa = dim(fit_aa$fit)

@@

### Cleaning RH curves
n_rh_curve_raw = 2400 # export from adaptive-amphistomy/r/07_process-licor.R
n_points_per_rh_curves_raw = 100 # export from adaptive-amphistomy/r/07_process-licor.R
ca_min = aa_args$ca_min
ca_max = aa_args$ca_max
n_min_points_per_rh_curve = aa_args$n_min_points_per_rh_curve
point_outlier_threshold = aa_args$point_outlier_threshold
aa_outlier_threshold = aa_args$aa_outlier_threshold
thinning_interval = aa_args$thinning_interval

@@

1. remove unreliable data points
2. remove hysteretic portion of \agcurve{} curves at low \gsw{}
3. remove outliers within each \agcurve{} curve
4. thin redundant data points within each \agcurve{} curve
5. remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves
6. Trim extreme \aax{} values

**Table showing number of curves and points before and after each cleaning step**

@@

*Procedure*: We fit provisional quadratic regressions for each curve using ordinary least squares with the `lm()` function in *R*. We sequentially removed data points with an absolute external studentized residual greater than `r point_outlier_threshold` until none remained.

#### Remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves

*Rationale*: We could not estimate \aax{} for replicates where amphi and pseudohypo \agcurve{} curves did not overlap.

*Procedure*: We removed replicates where the range of \gsw{} values for amphi and pseudohypo \agcurve{} curves did not overlap.

#### Thin redundant data points within each \agcurve{} curve

*Rationale*: Data points closely spaced along the \agcurve{} curve provide redundant information and may be highly correlated (i.e. pseudoreplication). This occurred because data was logged at a constant temporal interval, but the rate at which \gsw{} declined was not constant. Thinning reduces parameter estimation bias toward densely sampled regions of the curve which may not be the most biologically informative.

*Procedure*: We retained the maxima and minima \gsw{} for each curve and thinned all but one point per thinning interval of $`r thinning_interval`~\log(\si{\mol\raiseto{-2}\meter\raiseto{-1}\second})$, retaining the point nearest the midpoint of the interval.

#### Trim extreme \aax{} values

*Rationale*: Extreme \aax{} values were likely due to measurement error or leaf damage. Since amphi and pseudohypo \agcurve{} curves are measured on consecutive days, a poor calibration or a damaged leaf could cause a large difference in $A$ between days, which would appear as an extreme \aax{} value.

*Procedure*: We privionsally estimated \aax{} for each replicate by integrating over the range of \gsw{} values where amphi and pseudohypo \agcurve{} curves overlap. In this procedure, curve parameters were provisionally estimated using ordinary least squares with the `lm()` function in *R*. We then used point estimates of \aax{} for each replicate as the response variable in a linear model with light treatment, light intensity, accession, and all interactions as explanatory variables. This model was also fit using ordinary least squares with the `lm()` function in *R*. We classified  extreme \aax{} values as those with an absolute internal studentized residual $> `r aa_outlier_threshold`$. Because these values likely indicate significant measurement error or leaf damage, we removed \agcurve{} curves at both light intensities if either was classified as extreme.

### 2024-04-03 ms/ms.tex
finished drafting data cleaning methods

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  remove unreliable data points
\item
  remove hysteretic portion of \agcurve{} curves at low \gsw{}
\item
  remove outliers within each \agcurve{} curve
\item
  remove replicates with no overlap between amphi and pseudohypo
  \agcurve{} curves
\item
  remove replicates with outlier \aax values
\end{enumerate}

@@

\emph{Procedure}: For each

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  remove unreliable data points
\item
  remove hysteretic portion of \agcurve{} curves at low \gsw{}
\item
  remove outliers within each \agcurve{} curve
\item
  thin redundant data points within each \agcurve{} curve
\item
  remove replicates with no overlap between amphi and pseudohypo
  \agcurve{} curves
\item
  Trim extreme \aax{} values
\end{enumerate}

\textbf{Table showing number of curves and points before and after each
cleaning step}

@@

\emph{Procedure}: For each curve, we removed data points after \gsw{}
had reached its minimum unless there were fewer than 10 data points
remaining.

\paragraph{\texorpdfstring{Remove outliers within each \agcurve{}
curve}{Remove outliers within each  curve}}\label{remove-outliers-within-each-curve}

\emph{Rationale}: Individual outliers within \agcurve{} curves, usually
caused by transitory changes in chamber conditions, exert undue leverage
on parameter estimates and cause bias and/or low precision in parameter
estimates.

\emph{Procedure}: We fit provisional quadratic regressions for each
curve using ordinary least squares with the \texttt{lm()} function in
\emph{R}. We sequentially removed data points with an absolute external
studentized residual greater than until none remained.

\paragraph{\texorpdfstring{Remove replicates with no overlap between
amphi and pseudohypo \agcurve{}
curves}{Remove replicates with no overlap between amphi and pseudohypo  curves}}\label{remove-replicates-with-no-overlap-between-amphi-and-pseudohypo-curves}

\emph{Rationale}: We could not estimate \aax{} for replicates where
amphi and pseudohypo \agcurve{} curves did not overlap.

\emph{Procedure}: We removed replicates where the range of \gsw{} values
for amphi and pseudohypo \agcurve{} curves did not overlap.

\paragraph{\texorpdfstring{Thin redundant data points within each
\agcurve{}
curve}{Thin redundant data points within each  curve}}\label{thin-redundant-data-points-within-each-curve}

\emph{Rationale}: Data points closely spaced along the \agcurve{} curve
provide redundant information and may be highly correlated
(i.e.~pseudoreplication). This occurred because data was logged at a
constant temporal interval, but the rate at which \gsw{} declined was
not constant. Thinning reduces parameter estimation bias toward densely
sampled regions of the curve which may not be the most biologically
informative.

\emph{Procedure}: We retained the maxima and minima \gsw{} for each
curve and thinned all but one point per thinning interval of
\(0.05~\log(\si{\mol\raiseto{-2}\meter\raiseto{-1}\second})\), retaining
the point nearest the midpoint of the interval.

\paragraph{\texorpdfstring{Trim extreme \aax{}
values}{Trim extreme  values}}\label{trim-extreme-values}

\emph{Rationale}: Extreme \aax{} values were likely due to measurement
error or leaf damage. Since amphi and pseudohypo \agcurve{} curves are
measured on consecutive days, a poor calibration or a damaged leaf could
cause a large difference in \(A\) between days, which would appear as an
extreme \aax{} value.

\emph{Procedure}: We privionsally estimated \aax{} for each replicate by
integrating over the range of \gsw{} values where amphi and pseudohypo
\agcurve{} curves overlap. In this procedure, curve parameters were
provisionally estimated using ordinary least squares with the
\texttt{lm()} function in \emph{R}. We then used point estimates of
\aax{} for each replicate as the response variable in a linear model
with light treatment, light intensity, accession, and all interactions
as explanatory variables. This model was also fit using ordinary least
squares with the \texttt{lm()} function in \emph{R}. We classified
extreme \aax{} values as those with an absolute internal studentized
residual \(> 3.5\). Because these values likely indicate significant
measurement error or leaf damage, we removed \agcurve{} curves at both
light intensities if either was classified as extreme.
